# Reproducible Research: Peer Assessment 1
```{r}
library(lattice)
```
## Loading and preprocessing the data
Load csv data **activity.csv** and convert dates to **R Date class**  
```{r}
actdata <- read.csv("activity.csv")
actdata$date <- as.Date(actdata$date,"%Y-%m-%d")
```

## What is mean total number of steps taken per day?
Compute total number of steps per day  
```{r}
totsteps <- tapply(actdata$steps, actdata$date,sum)

```
Plot histogram of **total number of steps per day**
```{r}
hist(totsteps,col="blue",xlab="Total Steps per Day", 
      ylab="Frequency", main="Histogram of Total Steps taken per day")
```
Compute Mean total steps taken per day
```{r}
mean(totsteps,na.rm=TRUE)
```

Compute Median total steps taken per day
```{r}
median(totsteps,na.rm=TRUE)
```

## What is the average daily activity pattern?
Compute mean of steps over all days by time interval
```{r}
meansteps <- tapply(actdata$steps,actdata$interval,
                                 mean,na.rm=TRUE)
```
Timeseries plot of of the 5-minute interval and the average number of steps taken, averaged across all days
```{r}
plot(row.names(meansteps),meansteps,type="l",
     xlab="Time Intervals (5-minute)", 
     ylab="Mean number of steps taken (all Days)", 
     main="Average Steps Taken at 5 minute Intervals",
     col="blue")
```
Find the time interval that contains maximum average number of steps over all days
```{r}
interval_num <- which.max(meansteps)
interval_max_steps <- names(interval_num)
interval_max_steps
```
The ** `r  interval_max_steps `** minute  or ** `r  interval_num `th ** 5 minute interval contains the maximum number of steps on average across all the days


## Imputing missing values
Compute the number of NA values in the activity dataset
```{r}
num_na_values <- sum(is.na(actdata))
num_na_values 
```

Fill in missing values using the **average interval value across all days**
```{r}
na_indices <-  which(is.na(actdata))
imputed_values <- meansteps[as.character(actdata[na_indices,3])]
names(imputed_values) <- na_indices
for (i in na_indices) {
    actdata$steps[i] = imputed_values[as.character(i)]
}
sum(is.na(actdata)) 
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
days <- weekdays(actdata$date)
actdata$day_type <- ifelse(days == "Saturday" | days == "Sunday", 
                                "Weekend", "Weekday")
meansteps <- aggregate(actdata$steps,
                                    by=list(actdata$interval,
                                            actdata$day_type),mean)
names(meansteps) <- c("interval","day_type","steps")
xyplot(steps~interval | day_type, meansteps,type="l",
       layout=c(1,2),xlab="Interval",ylab = "Number of steps")

```
Let's compute the mean, median, max and min of the steps across all intervals and days by Weekdays/Weekends
```{r}
tapply(meansteps$steps,meansteps$day_type,
       function (x) { c(MINIMUM=min(x),MEAN=mean(x),
                        MEDIAN=median(x),MAXIMUM=max(x))})
```